\chapter{Conclusion and Evaluation}

So far in this thesis many topics have been covered. Chapter \ref{theory} coveres theory related to snow plowing and route optimization problems. Then in Chapter \ref{architecture_and_implementation} the implementation of the MA is discussed. In chapters \ref{cha:evolutionary_algorithm_configuration} and \ref{cha:experiments_and_results} the performance and outputs of the MA are covered.

This chapter will be focused on rounding off the thesis. It will start off by evaluating what questions have been answered and to what extent. Then there will be a section on the contributions to the field of the work. Finally the future work that can be done to expand upon what has been covered in this thesis will be outlined.

\section{Conclusion}

When evaluating the work in this thesis, the research questions should be considered. The first research question is stated in Chapter \ref{cha:introduction} as \emph{\enquote{Obtain and present the related work that gives the context for this thesis}}. It is subsequently treated in Chapter \ref{theory}, which covered both the background for snow plowing and the theory surrounding route optimization problems. Chapter \ref{sec:routing_algorithms} discusses why a MA is chosen as the approach to the route optimization problem, and later in Chapter \ref{sec:evolutionary_algorithms} MA's are explained more in details. Then in Chapter \ref{sub:fitness_for_snow_plowing} the theory on how to adapt a MA better to the task of snow plowing is dealt with. Finally in chapters \ref{sub:all_pairs_shortest_path} and \ref{sub:the_split_algorithm} the reasoning behind the adoption of a couple of known algorithm's is laid out. Thus Chapter \ref{theory} presents related work, ties it to the setting of the thesis, and thereby answeres RQ1.

The second research question in this thesis is given as \emph{\enquote{Configure the memetic algorithm so that it performs well}}. To do the configuration a test case that is simple for a human to verify the correctness of the solution to is made and presented in Chapter \ref{cha:evolutionary_algorithm_configuration}. The MA is then applied to the problem with a variety of configurations, and a combination of parameters that work well is found. One might think that at this point RQ2 has been completely answered, but when treating the results in Chapter \ref{cha:evolutionary_algorithm_configuration} an interesting pattern in the MAs solutions becomes apparent. The MA, while capable of finding good solutions, has not yet found an optimal solution, even in cases where it is trivial for humans to do so. However, the results it produced are of a sufficient quality to move forwards. This is mainly because the routes found by the MA in Chapter \ref{sec:map_data_from_trondheim} are better than the ones currently being driven. The MA can thus be considered to perform well, and RQ2 satisfyingly addressed by the work presented in Chapter \ref{cha:evolutionary_algorithm_configuration}.

Chapter \ref{sec:map_data_from_trondheim} is also relevant when evaluating how the third research question, \emph{\enquote{Determine whether the memetic algorithm independently can find a route in Trondheim optimized for length weighted by speed with a random starting point}}, is dealt with. There it is showed what results are obtained when the MA is applied to data from Trondheim. Moreover, when looking at the outcomes, it would seem that the MA has indeed managed to produce routes that are good in terms of length weighted by speed and with random starting points. However, for the routes generated in Chapter \ref{sec:map_data_from_trondheim} it can be argued that they do not have a random starting point because it is chosen to be the first required road piece in a route that appears in the input file to the MA. Yet it should be considered that which road piece is the first is rather arbitrary, and that the route is a giant circle and starting on a required point on that circle. Then it becomes clear that it does not really matter where the driver choses to start as long as he drives the shortest possible path to the route, and that the starting point can be randomly chosen and the route be exactly the same. In that light RQ3 has been fully answered by the work presented in Chapter \ref{sec:map_data_from_trondheim}. It could also be argued that it shows that given a model of sufficiently high quality, route optimization for snow plowing in Trondheim, or anywhere else in Norway with similar constraints, is possible and feasible.

The question that remains at the end then is whether the goal of \emph{\enquote{Make a system that can generate optimal routes for snow plowing in Trondheim}} has been reached. Because the work in Chapter \ref{cha:evolutionary_algorithm_configuration} relating to RQ2 shows that the MA has not found optimal solutions, the answer has to be no. However, the work still has merits in terms of that the solutions it finds are estimated to be of a high quality by the domain experts given the input used. As such the project should be considered a success even though its goal has not been reached.

% \section{Discussion}

\section{Contributions}

The work carried out in this thesis does not represent an improvement upon know solutions to rotue optimization problems, nor does it introduce any new techniques for solving them. Therefore, it does not constitute a contribution to pushing the current state of the art of the research further. However it does explore how solutions to route optimization for snow plowing in Trondheim can be done, and in the process it has verified that MAs can be applied with success to real-world route optimization problems. Additionally it has been shown that the routes can be compared to those being driven by humans in a meaningful way. The most significant outcome of the work for the field is that it shows that the map data required to make routes for snow plowing in Norway exists, is publicly available, and can successfully be used for route optimization.

\section{Future Work}

First and foremost it should be examined why the MA is incapable of finding the optimal solution even in simple cases. Even if it happens to take an unreasonably long time, it would be nice to know that the MA is capable of finding the optimal solution if run indefinitely. On a related note, the overall performance of the MA should be improved. As the implementation is now, it only works serially on one processor core. It could benefit greatly by being optimized for running in parallel on several cores, or perhaps even on a graphics card. The creation of each child, and evaluation of its fitness could be done parallelly, greatly increasing how fast each generation is processed and the overall performance of the MA.

Another improvement that should be done is to merge as many road pieces in the underlying data before it is fed to the MA, or drawn on the map. When merging them before feeding the data to the MA the number of elements in the network to process is reduced, and the performance will improve. The reasons for the improvement would be amongst other things that there would be fewer elements to iterate over each time the fitness is calculated, and when there are fewer pieces of road that are required the number of possible ways to make a route is reduced dramatically. It is also desirable to reduce the number of road pieces when drawing up the generated routes as maps. It would make the routes easier to read for the drivers because it would remove much detail that is not necessary for their work.

In relation to how the maps are read, a nice expansion upon the work would be to make an interactive map of the routes that does not require proficiency in setting up or using QGIS. If the maps of the routes are available, for instance, in an online format on a webpage, it would be possible for the drivers or other interested parties to view larger routes in more detail. It could also make it easier to compare routes.

However, other than pure technical improvements in how the routes are generated and represented, there are also improvements that can be made to the experiments, and also the input data itself. 
For instance, considering the discovery about population sizes in Chapter Chapter \ref{sec:results}. There the effects of combinations of parameters reveal that in some cases a smaller population size yields better reseults than a bigger one. It is clear that it should be examined further to obtain a certain proof of that the only reason for the odd result that given much randomness a smaller population is better if you reinsert the best individual of the previous generation each generation.

Improvements that can be made to the heuristic are however maybe the most important for increasing the quality of the routes. First and foremost by using more of the available features of the underlying data and working them into the data that is sent to the MA. If the width of the road and the number of files for instance were included in the input data the heuristic could be more precise in determining the amount of vehicles and passes needed for clearing a given road. Coupled with data about what types of vehicles are available and how many it could be used to determine an optimal sectoring of the road network.

With more data available, the heuristic could also be made to optimize other things than just length, or a prioritized combination of features of the routes to optimize. Considering the desire to get the traffic running as smoothly as fast as possible, optimizing for clearing the most heavily trafficked roads first and finishing the job as quickly as possible could be one thing to gear the MA towards. Another thing that could be interesting to optimize for is the fuel consumption of the plowing. This would affect the environmental impact of the plowing, but would also be important in reducing the total cost of the plowing operations.

% Reduce constant factors (of running time (ex some loops could be n\^2 instead of 2\*n\^2)).

% Take todays weather into consideration (i.e., make a real-time system, and differentiate between emergency servicing and maintenance work).

% Give output in terms of not only what we currently do, but also total distance, amount of fuel spent, time spent, vehicle durability (how much is their lifespan reduced by by doing the suggested route), or environmental impact. This will allow the operator to choose what to optimize, and it could even be a weighting among all the different evaluation types. Input could then for instance be fuel > time > distance > environmental impact, or 90\% environment, 80\% fuel, 85\%time, etc.

\cleardoublepage
